---
title: "Reproducible Research: Peer Graded Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

```{r}

library(dplyr)
library(ggplot2)

dfActivity <- read.csv("activity.csv",header=TRUE)
dfActivity$date <- as.Date(dfActivity$date)
dfActivity$steps <- as.numeric(dfActivity$steps)
dfActivity$interval <- as.numeric(dfActivity$interval)

dfActivity.steps <- dfActivity %>% group_by(date) %>% summarise(steps=sum(steps,na.rm = T))
```

## What is mean total number of steps taken per day?
```{r Figure1}
ggplot(dfActivity.steps, aes(x=steps)) +  geom_histogram(binwidth=2000,fill="orange",color="white")+
  ylim(0,15)
```

```{r}
dfActivity %>% summarise(Mean=mean(dfActivity$steps,na.rm = T),Median=median(dfActivity$steps,na.rm = T))
```

## What is the average daily activity pattern?
```{r}
dfActivity.avgsteps <- dfActivity %>% group_by(interval) %>% summarise(steps=mean(steps,na.rm = T))
```

```{r Figure2}
ggplot(dfActivity.avgsteps, aes(x=interval, y=steps,col="red")) + geom_line()
```

## The 5-minute interval that, on average, contains the maximum number of steps
```{r}
dfActivity.avgsteps[which.max(dfActivity.avgsteps$steps), ]
```

## Imputing missing values
### Total number of missing values in the dataset
```{r}
sum(is.na(dfActivity))
```

### Replace the missing values in the dataset with the Mean
```{r}
dfActivity$stepsnew <- dfActivity$steps
dfActivity$stepsnew[is.na(dfActivity$stepsnew)] <- mean(dfActivity$steps,na.rm = T)
```

### Histogram for the New dataset with values replaced.
```{r}
par(mfrow = c(1,2))
dfActivity.stepsNew <- dfActivity %>% group_by(date) %>% summarise(steps=sum(stepsnew))
```

```{r Figure3}
par(mfrow = c(1,2))
hist(dfActivity.steps$steps,xlab="Total No. of Steps",ylab="No. of Days",main="NA's Omitted",col="chocolate")
hist(dfActivity.stepsNew$steps,xlab="Total No. of Steps",ylab="No. of Days",main="NA's Replaced",col="magenta")
```

## Are there differences in activity patterns between weekdays and weekends?
```{r}
weekdays <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
dfActivity$wday <- factor(weekdays(as.Date(dfActivity$date)) %in% weekdays, levels=c(FALSE, TRUE),labels=c('weekend', 'weekday'))
dfActivity.wday <- dfActivity %>% group_by(interval,wday) %>% summarise(steps=mean(steps,na.rm = T))
```

```{r Figure4}
ggplot(dfActivity.wday, aes(x=interval, y=steps))+ geom_line(color="blue")+facet_wrap(~dfActivity.wday$wday,  ncol=1)
```

